*flippant* -- An R package for the automated analysis of dithionite scramblase

### 2015-10-16 inst/manuscript/manuscript.rmd
Integrate usage of "splitByExperiment" functonality;

figure1 <- multipanelfigure(
  widths = rep(17.22/3-2*0.25, 3), 
  heights = rep(18/5-4*0.25, 5), 
  units = "cm",
  interPanelSpacing = 0.25)

knitr::opts_chunk$set(
  fig.width = 6.78, 
  fig.height = 7.09, 
  warning = FALSE, 
  message = FALSE)
```

@@

allows for the complete generation of the PPR plot in publication quality from 
the user assembled spreadsheet of input data (`x`), including automatically 
added graphical representations the PPR value equal to the fit constant 
$\alpha$ ($P_{\geq 1} = 1 - e^{-1}$), which implies all liposomes on average 
carrying one scramblase and 63% have one or more 
[i.e. are active; @goren_constitutive_2014]. The behavior of the function may be
adjusted to use scramblase-mediated fluorescence reduction normalized to either
the maximum measurement in the series (`scaleTo = "data"`) or the plateau of an
exponential fit to it (`scaleTo = "model"`; for reasoning see above) and whether
or not to require the exponential curve fitted to the data to go through the
origin.

@@

```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.cap = "***flippant*-based reanalysis of data in Figure 4C/D from @ploier_perturbed_submitted.** (A) represents spectral traces as produced by a call to the function `scramblaseAssayTraces`, while (B) shows protein to phospholipid ration (PPR) plots for the data using a x-axis trimmed to a maximum value of 2 (see text for details)."}
setwd(tmpDir)
figure1 <- addpanel(
  figure = figure1,
  topPanel = 1,
  bottomPanel = 3,
  leftPanel = 1, 
  rightPanel = 3, 
  panel = scramblaseAssayTraces("inputTable.txt", timeMaxSec = 350) + theme_bw())
figure1 <- addpanel(
  figure = figure1,
  topPanel = 4,
  bottomPanel = 5,
  leftPanel = 1, 
  rightPanel = 3, 
  panel = scramblaseAssayPlot("inputTable.txt") + 
    theme_bw() + 
    xlim(0, 2) + 
    xlab(expression("PPR"~~("mg"/"mmol"))) +
    ylab(expression(P[{phantom()>=1}])))
grid.draw(figure1)
```

@@

```{r, eval = FALSE}
# Plot PPR plot (Figure 1B)
scramblaseAssayPlot("inputTable.txt") + xlim(0, 2)
# Write out results in tabular manner
scramblaseAssayStats("inputTable.txt")
```

@@

The first call results in the protein to phospholipid ratio (PPR) plot shown in
Figure 1B. Using the customization facilities provided by `ggplot2` 
[@wickham_ggplot2_2009] as a plotting system the plot's x-axis is trimmed to
avoid visual compression of all data by a single data point from the wild type 
rhodopsin series in experiment 1, where $\text{PPR}=`r max_PPR`$. Dashed lines
are added automatically as graphical representations of the fit constant 
$\alpha$, but are difficult to evaluate visually.

Recapitulation of the calculations underlying PPR plot generation by 
`scramblaseAssayPlot` in a call to `scramblaseAssayStats` produces tabular 
output represented in Table 1, which includes series-specific fit constants 
$\alpha$ in numeric and easily accessible form.

```{r, echo = FALSE}
setwd(tmpDir)
ppr_results <- scramblaseAssayStats("inputTable.txt")
ppr_results_alpha_average <- ppr_results %>%
  split(ppr_results$`Experimental Series`) %>% 
  sapply(function(y){mean(y$`Fit Constant`, na.rm = TRUE)}) %>%
  round(2)
ppr_results_alpha_average_names <- names(ppr_results_alpha_average) %>%
  head(n = length(ppr_results_alpha_average) - 1) %>%
  paste(collapse = ", ") %>%
  paste("and") %>%
  paste(tail(names(ppr_results_alpha_average), n = 1))
ppr_results_alpha_average_print <- ppr_results_alpha_average %>%
  unname() %>%
  head(n = length(ppr_results_alpha_average) - 1) %>%
  paste(collapse = ", ") %>%
  paste("and") %>%
  paste(tail(ppr_results_alpha_average, n = 1))
knitr::kable(
  ppr_results,
  row.names = FALSE,
  caption = "**Tabular output of *flippant*-based reanalysis of data in Figure 4C/D from @ploier_perturbed_submitted.** An plain text-version of this output is produced by calling `scramblaseAssayStats`.")
```

Mean values of fit constant $\alpha$ correspond with 
`r ppr_results_alpha_average_print` for `r ppr_results_alpha_average_names`, 
respectively, very well to the prior manual analysis by 
@ploier_perturbed_submitted, where $\alpha \approx 0.13$ was reported for mutant 
and $\alpha \approx 0.26$ for wild type rhodopsin. Based on the manual analysis 
of the data we conclude in @ploier_perturbed_submitted that the mutant 
rhodopsins investigated are reconstituted into proteoliposomes in monomeric 
form, contrary to the wild type, which is reconstituting as a homodimer. 
*flippant* analysis of the data shows identical trends and close matches of the
results, supporting identical conclusions.

@@

```{r, echo = FALSE, cache = TRUE}
setwd(tmpDir)
analysis_timing_n <- 100
analysis_timing_function <- function(){
  scramblaseAssayTraces(
    "inputTable.txt",
    timeMaxSec = 350)
  scramblaseAssayPlot("inputTable.txt")
  scramblaseAssayStats("inputTable.txt")
  gc()
}
analysis_timing <- microbenchmark(
  analysis_timing_function(),
  times = analysis_timing_n)

>>>>

figure1 <- multipanelfigure(
  widths = rep(17.22/3-2*0.25, 3), 
  heights = rep(27/7-5*0.25, 7), 
  units = "cm",
  interPanelSpacing = 0.25)

knitr::opts_chunk$set(
  fig.width = 6.78, 
  fig.height = 10.63, 
  warning = FALSE, 
  message = FALSE)
```

@@

allows for the complete generation of the PPR plot in publication quality from 
the user assembled spreadsheet of input data (`x`), including automatically 
added graphical representations the PPR value equal to the fit constant 
$\alpha$ ($P_{\geq 1} = 1 - e^{-1}$), which implies all liposomes on average 
carrying one scramblase and 63% have one or more 
[i.e. are active; @goren_constitutive_2014]. The behavior of the function may be
adjusted to use scramblase-mediated fluorescence reduction normalized to either
the maximum measurement in the series (`scaleTo = "data"`) or the plateau of an
exponential fit to it (`scaleTo = "model"`; for reasoning see above) and whether
or not to require the exponential curve fitted to the data to go through the
origin. The `splitByExperiment` argument is used to indicate whether independent
experiments are integrated into a single analysis (`splitByExperiment = FALSE`)
or treated separately (`splitByExperiment = TRUE`; see below for example usage).

@@

```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.cap = "***flippant*-based reanalysis of data in Figure 4C/D from @ploier_perturbed_submitted.** (A) represents spectral traces as produced by a call to the function `scramblaseAssayTraces`, while (B) and (C) show protein to phospholipid ration (PPR) plots for the data using a x-axis trimmed to a maximum value of 2 (see text for details). In (B) data is separately plotted for independent experiments, while (C) combines all dat points, aiming for reliability of fit."}
setwd(tmpDir)
figure1 <- addpanel(
  figure = figure1,
  topPanel = 1,
  bottomPanel = 3,
  leftPanel = 1, 
  rightPanel = 3, 
  panel = scramblaseAssayTraces("inputTable.txt", timeMaxSec = 350) + theme_bw())
figure1 <- addpanel(
  figure = figure1,
  topPanel = 4,
  bottomPanel = 5,
  leftPanel = 1, 
  rightPanel = 3, 
  panel = scramblaseAssayPlot("inputTable.txt") + 
    theme_bw() + 
    xlim(0, 2) + 
    xlab(expression("PPR"~~("mg"/"mmol"))) +
    ylab(expression(P[{phantom()>=1}])))
figure1 <- addpanel(
  figure = figure1,
  topPanel = 6,
  bottomPanel = 7,
  leftPanel = 1,
  rightPanel = 2,
  panel = scramblaseAssayPlot("inputTable.txt", splitByExperiment = FALSE) + 
    theme_bw() + 
    xlim(0, 2) + 
    xlab(expression("PPR"~~("mg"/"mmol"))) +
    ylab(expression(P[{phantom()>=1}])))
grid.draw(figure1)
```

@@

```{r, eval = FALSE}
# Plot PPR plot (Figure 1B)
scramblaseAssayPlot("inputTable.txt") + xlim(0, 2)
```

@@

The call results in the protein to phospholipid ratio (PPR) plot shown in
Figure 1B. Using the customization facilities provided by `ggplot2` 
[@wickham_ggplot2_2009] as a plotting system the plot's x-axis is trimmed to
avoid visual compression of all data by a single data point from the wild type 
rhodopsin series in experiment 1, where $\text{PPR}=`r max_PPR`$. Dashed lines
are added automatically as graphical representations of the fit constant 
$\alpha$.

Figure 1B emphasizes reproducebility between independent experiments
("Experiment 1" and "Experiment 2") on the four probed rhodopsins. Forgoing
reproducibility analysis in favor of a potentially more reliable result and
following more closely the analysis in Figure 4C/D by
@ploier_perturbed_submitted, a single fit may be performed for the combined data
from all experiments as follows, resulting in Figure 1C:

```{r, eval = FALSE}
# Plot PPR plot (Figure 1C)
scramblaseAssayPlot("inputTable.txt", splitByExperiment = FALSE) + xlim(0, 2)
```

The graphical representation of the fit constant $\alpha$ in the PPR plots is
difficult to visually evaluate and easy access to the numeric values desireable.
Recapitulation of the calculations underlying PPR plot generation for Figure 1C
by `scramblaseAssayPlot` in a call to `scramblaseAssayStats`, this is achieved
and produces tabular output represented in Table 1.

```{r, eval = FALSE}
# Write out results in tabular manner (Table 1)
scramblaseAssayStats("inputTable.txt" splitByExperiment = FALSE)
```

```{r, echo = FALSE}
setwd(tmpDir)
ppr_results <- scramblaseAssayStats("inputTable.txt", splitByExperiment = FALSE)
ppr_results_names <- ppr_results$`Experimental Series` %>%
  head(n = nrow(ppr_results) - 1) %>%
  paste(collapse = ", ") %>%
  paste("and") %>%
  paste(tail(ppr_results$`Experimental Series`, n = 1))
ppr_results_alpha_print <- ppr_results$`Fit Constant` %>%
  round(2) %>%
  head(n = nrow(ppr_results) - 1) %>%
  paste(collapse = ", ") %>%
  paste("and") %>%
  paste(tail(ppr_results$`Fit Constant`, n = 1))
knitr::kable(
  ppr_results,
  row.names = FALSE,
  caption = "**Tabular output of *flippant*-based reanalysis of data in Figure 4C/D from @ploier_perturbed_submitted.** An plain text-version of this output is produced by calling `scramblaseAssayStats`.")
```

The resulting values for the fit constant $\alpha$ correspond with
`r ppr_results_alpha_print` for `r ppr_results_names`, respectively, very well
to the prior manual analysis by @ploier_perturbed_submitted, where
$\alpha \approx 0.13$ was reported for mutant and $\alpha \approx 0.26$ for wild
type rhodopsin. Based on the manual analysis of the data we conclude in @ploier_perturbed_submitted that the mutant rhodopsins investigated are
reconstituted into proteoliposomes in monomeric form, contrary to the wild type,
which is reconstituting as a homodimer. 
*flippant* analysis of the data shows identical trends and close matches of the
results, supporting identical conclusions.

@@

```{r, echo = FALSE, cache = TRUE}
setwd(tmpDir)

analysis_timing_n <- 100

analysis_timing_function <- function(){
  scramblaseAssayTraces(
    "inputTable.txt",
    timeMaxSec = 350)
  scramblaseAssayPlot("inputTable.txt")
  scramblaseAssayPlot("inputTable.txt", splitByExperiment = FALSE)
  scramblaseAssayStats("inputTable.txt", splitByExperiment = FALSE)
}

analysis_timing <- microbenchmark(
  analysis_timing_function(),
  times = analysis_timing_n)
